### 2017-04-05 writing/paper/sections/meshpatterns.tex
Comments from Henning up to section 4

A \emph{permutation} is a bijection from the set \(\nrange{n} = \setrange{1}{n}\) to itself. The
set of all such bijections on this set if denoted \(\mathfrak{S}_n\) and has size \(n!\). We can
denote an individual permutation \(\pi\in\mathfrak{S}_n\) in \emph{one-line notation} by writing the
entries of the permutation in order, therefore \(\pi = \perm{\pi(1),\pi(2),\dotso,\pi(n)}\). The
set \(\mathfrak{S}_n\) has exactly one element, the empty permutation \(\varepsilon\).

\begin{definition}{(Order isomorphism.)}
    Two substrings \(\alpha_1\alpha_2\dotsm\alpha_n\) and
    \(\beta_1\beta_2\dotsm\beta_n\) are said to be \emph{order isomorphic}
    if they share the same relative order, \ie, \(\alpha_r<\alpha_s\) if and
    only if \(\beta_r<\beta_s\).
\end{definition}

The definition of order isomorphism allows us to give the meaning of containment for classical
permutation patterns.
\begin{definition}
    A permutation \(\pi\in\mathfrak{S}_n\) contains the pattern \(\sigma\in\mathfrak{S}_k\) (\mbox{denoted} 
    \(\sigma \preceq \pi\)) if there is some sequence\( i_1,i_2,\dotsm{},i_k\)
    such that \(1\le i_1<i_2<\dotsm<i_k\le n\) and the sequence \(\pi(i_1)\pi(i_2)\dotsm\pi(i_k)\) is order isomorphic to
    \(\sigma(1)\sigma(2)\dotsm\sigma(k)\). If \(\pi\) does not contain \(\sigma\), we say that 
    \(\pi\) \emph{avoids} \(\sigma\).
\end{definition}
\begin{example}
\label{ex:contexmpl}
The permutation \(\pi = \perm{2,4,1,5,3}\) contains the pattern 
\(\sigma = \perm{2,3,1}\), since the second, fourth and fifth elements (\(453\)) are order 
isomorphic to \(\perm{231}\), it also contains the occurrence \(241\) of the sam pattern. The permutation 
\(\perm{2,4,1,5,3}\) avoids the pattern \(\perm{3,2,1}\).
\end{example}

We denote the set of permutations of length \(n\) avoiding a pattern \(\sigma\) as \(\Av_n(\sigma)\)
and \(\av{\sigma}=\bigcup_{i=0}^{\infty}\Av_i(\sigma)\).

We can display a permutation graphically in a \emph{plot}, in such a plot we display the points
\(G(\pi) = \{(i,\pi(i))\mid i\in[1,n]\}\) in a Cartesian coordinate system. The plots of the permutations \(\pi = 
\perm{2,4,1,5,3}\) and \(\sigma = \perm{2,3,1}\) can be seen in \FigureRef{fig:plots}.
\FigureRef{fig:containment} shows the containment of \(\sigma\) in \(\pi\) as in \ExampleRef{ex:contexmpl}.

@@

Formally defined by \textcite{journals/combinatorics/BrandenC11},
an \emph{occurrence} of \(p\) in \(\pi\) is a subset \(\omega\) of the plot of
\(\pi, G(\pi) = \{(i,\pi(i)\mid i\in[1,n]\}\) such that there are order-preserving
injections \(\alpha,\beta:[1,k]\mapsto[1,n]\) satisfying the following two conditions.

\noindent Firstly, \(\omega\) is an occurrence of \(\tau\) in the classical sense
  \begin{enumerate}[i.]
    \item \(\omega = \{(\alpha(i),\beta(j)):(i,j)\in G(\tau)\}\)
  \end{enumerate}\saveenum
  Define \(R_{ij} = [\alpha(i)+1,\alpha(i+1)-1]\times[\beta(j)+1,\beta(j+1)-1]\)
   for \(i,j\in[0,k]\) where \(\alpha(0)=\beta(0)=0\) and \(\alpha(k+1)=\beta(k+1)=n+1\). Then
   the second condition is
  \begin{enumerate}[i.]\resetenum
  \item if \(\boks{i,j} \in R \text{ then } R_{ij} \cap G(\pi) = \emptyset\)
\end{enumerate}
We call \(R_{ij}\) the region corresponding to \(\boks{i,j}\).
We define containment of a mesh pattern \(p\) in another mesh pattern \(q\) as above,
with the additional condition that if \(\boks{i,j}\in R\text{ then } R_{ij}\) is
contained in the mesh set of \(q\), in this case we call \(p\) a \emph{subpattern}
of \(q\).
\begin{definition}
A mesh pattern \(q=(\kappa,T)\) \emph{contains} a mesh pattern \(p=(\tau,R)\) as a \emph{subpattern}
if \(\kappa\) contains \(\tau\) and \(\left(\bigcup_{\boks{i,j}\in R} R_{ij}\right)\subseteq T\).
\end{definition}

@@

Let us consider the plot for the permutation \(\pi\). The subsequence \(325\) is
an occurrence of \(213\) in the classical sense and the remaining points of \(\pi\)
are not contained in the regions corresponding to the shaded boxes in \(p\).
\begin{equation*}
    \pattern{1,3,5}{3,4,2,1,5}{0/2,0/3,0/4,
                               1/0,1/1,1/2,
                               2/0,2/1,2/2,
                               3/2,3/3,3/4,
                               4/2,4/3,4/4}
\end{equation*}
The subsequence \(325\) is therefore an occurrence of the pattern \(p\) in \(\pi\) and
\(\pi\) contains \(p\).

>>>>

A \emph{permutation} is a bijection from the set \(\nrange{n} = \setrange{1}{n}\) to itself. The
set of all such bijections is denoted \(\mathfrak{S}_n\) and has \(n!\) elements. We can
denote an individual permutation \(\pi\in\mathfrak{S}_n\) in \emph{one-line notation} by writing the
entries of the permutation in order, therefore \(\pi = \perm{\pi(1),\pi(2),\dotso,\pi(n)}\). The
set \(\mathfrak{S}_0\) has exactly one element, the empty permutation \(\varepsilon\).

\begin{definition}{(Order isomorphism.)}
    Two strings \(\alpha_1\alpha_2\dotsm\alpha_n\) and
    \(\beta_1\beta_2\dotsm\beta_n\) are said to be \emph{order isomorphic}
    if they share the same relative order, \ie, \(\alpha_r<\alpha_s\) if and
    only if \(\beta_r<\beta_s\).
\end{definition}

The definition of order isomorphism allows us to give the meaning of containment for classical
permutation patterns.
\begin{definition}
    A permutation \(\pi\in\mathfrak{S}_n\) \emph{contains} the permutation \(\sigma\in\mathfrak{S}_k\) (\mbox{denoted}
    \(\sigma \preceq \pi\)) if there is some sequence \( i_1,i_2,\dotsm{},i_k\)
    such that \(1\le i_1<i_2<\dotsm<i_k\le n\) and the sequence \(\pi(i_1)\pi(i_2)\dotsm\pi(i_k)\) is order isomorphic to
    \(\sigma(1)\sigma(2)\dotsm\sigma(k)\). If \(\pi\) does not contain \(\sigma\), we say that
    \(\pi\) \emph{avoids} \(\sigma\). If this is the case the sequence
    \(\pi(i_1)\pi(i_2)\dotsm\pi(i_k)\) is called an \emph{occurrence} of \(\sigma\)
    in \(\pi\). In this context \(\sigma\) is called a (\emph{classical})
    \emph{permutation pattern}.
\end{definition}

\begin{example}
\label{ex:contexmpl}
The permutation \(\pi = \perm{2,4,1,5,3}\) contains the pattern
\(\sigma = \perm{2,3,1}\), since the second, fourth and fifth elements (\(453\)) are order
isomorphic to \(\perm{231}\). The permutation also contains the occurrence \(241\) of the same pattern. The permutation
\(\perm{2,4,1,5,3}\) avoids the pattern \(\perm{3,2,1}\).
\end{example}

We denote the set of permutations of length \(n\) avoiding a pattern \(\sigma\) as \(\Av_n(\sigma)\)
and let \(\av{\sigma}=\bigcup_{i=0}^{\infty}\Av_i(\sigma)\).

We can display a permutation graphically in a \emph{plot}, where we display the points
\(G(\pi) = \{(i,\pi(i))\mid i \in \nrange{n} \}\) in a Cartesian coordinate system. The plots of the permutations \(\pi =
\perm{2,4,1,5,3}\) and \(\sigma = \perm{2,3,1}\) can be seen in \FigureRef{fig:plots}.
\FigureRef{fig:containment} shows the containment of \(\sigma\) in \(\pi\) as in \ExampleRef{ex:contexmpl}.

@@

Formally defined by \textcite{journals/combinatorics/BrandenC11},
an \emph{occurrence} of \(p\) in \(\pi\) is a subset \(\omega\) of the plot of
\(\pi, G(\pi) = \{(i,\pi(i)\mid i\in\nrange{n}\}\) such that there are order-preserving
injections \(\alpha,\beta:\nrange{k}\mapsto\nrange{n}\) satisfying the following two conditions.

\noindent Firstly, \(\omega\) is an occurrence of \(\tau\) in the classical sense
  \begin{enumerate}[i.]
    \item \(\omega = \{(\alpha(i),\beta(j)):(i,j)\in G(\tau)\}\).
  \end{enumerate}\saveenum
  Define \(R_{ij} = [\alpha(i)+1,\alpha(i+1)-1]\times[\beta(j)+1,\beta(j+1)-1]\)
   for \(i,j\in[0,k]\) where \(\alpha(0)=\beta(0)=0\) and \(\alpha(k+1)=\beta(k+1)=n+1\). Then
   the second condition is
  \begin{enumerate}[i.]\resetenum
  \item if \(\boks{i,j} \in R \text{ then } R_{ij} \cap G(\pi) = \emptyset\).
\end{enumerate}
We call \(R_{ij}\) the \emph{region corresponding to} \(\boks{i,j}\).

@@

Let us consider the plot for the permutation \(\pi\). The subsequence \(325\) is
an occurrence of \(213\) in the classical sense and the remaining points of \(\pi\)
are not contained in the regions corresponding to the shaded boxes in \(p\).
\begin{equation*}
    \pattern{1,3,5}{3,4,2,1,5}{0/2,0/3,0/4,
                               1/0,1/1,1/2,
                               2/0,2/1,2/2,
                               3/2,3/3,3/4,
                               4/2,4/3,4/4}
\end{equation*}
The subsequence \(325\) is therefore an occurrence of the pattern \(p\) in \(\pi\).

We define containment of a mesh pattern \(p\) in another mesh pattern \(q\) as above,
with the additional condition that if \(\boks{i,j}\in R\text{ then } R_{ij}\) is
contained in the mesh set of \(q\).

\begin{definition}
A mesh pattern \(q=(\kappa,T)\) \emph{contains} a mesh pattern \(p=(\tau,R)\) as a \emph{subpattern}
if \(\kappa\) contains \(p\) and \(\left(\bigcup_{\boks{i,j}\in R} R_{ij}\right)\subseteq T\).
\end{definition}

TODO: I changed a symbol in the above definition. See the comment.
% if \(\kappa\) contains \(\tau\)
% to
% if \(\kappa\) contains \(p\)
% that is what we want right?

TODO: Change the following example to an example of a mesh pattern in a mesh
pattern.
